---
title: "AirQuality Dataset Exploratoration"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: yeti
    code_folding: hide
css: styles.css
---
<style>
  h1 {
     font-weight: bold;
  }
</style>

------------------------------------------------------------------------


[The air quality index (AQ)](https://data.cityofnewyork.us/Environment/Air-Quality/c3uy-2p5r) originated from NYC Open Data, contributed by the Department of Hygiene and Mental Health and include air quality indexes measured across boroughs in NYC.

In this webpage we will explore the AQI variable.

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

library(tidyverse)
library(ggplot2)
library(leaflet)

knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE)
```
  &nbsp;

## Load Data

```{r}
#import data
air_quality_df = read_csv("../data/air_quality/Air_Quality_20231126.csv") |>
  janitor::clean_names() |>
  mutate(
    start_date = mdy(start_date),
    year = year(start_date)
  ) |>
  filter(year == "2019")

joined_uhf_34_42 =  read_csv("../data/geocoding/joined_uhf_34_42.csv")

# add zip codes
air_geo =
  air_quality_df |>
  mutate(
    uhf34 = case_when(geo_type_name == "UHF34" ~ "1")
  )

#break up by geocoding system, add zipcodes

# uhf34
air_zip_neighborhoods_34 =
  air_geo |>
  filter(uhf34 == "1") |>
  select(-uhf34, - message)

air_zip_neighborhoods_34 =
  left_join(air_zip_neighborhoods_34, y = (joined_uhf_34_42),
            by = join_by("geo_join_id" == "uhf34")) 

air_final =
  air_zip_neighborhoods_34 |>
  select(-uhf34_neighborhood) |>
filter(!duplicated(unique_id))


head(air_final) |>
  knitr::kable()

```

```{r, eval = FALSE}
# save output for ease in other files
write_csv(air_zip_uhf, here("data/air_quality/air_zip_uhf.csv"))
```

 &nbsp;
  
## Summary Statistics  Across 2019

Average Air Quality Index Across Time Periods in 2019

```{r}
annual_aq = air_final |>
  filter(time_period == "Annual Average 2019") 

annual_aq|>
  ggplot(aes(x = time_period, y = data_value)) +
  geom_boxplot(fill = "skyblue", color = "black", alpha = 0.7) +
  labs(
       x = "Time Period",
       y = "Air Quality Index") +
  theme_minimal()

# Calculate statistics
midpoint <- median(annual_aq$data_value)
low_quantile <- quantile(annual_aq$data_value, 0.25)
top_quantile <- quantile(annual_aq$data_value, 0.75)

```

Summary of data particles: Fine particles (PM 2.5), Nitrogen dioxide (NO2), and Ozone (O3)

```{r}
air_final |>
  group_by(name) |>
  summarise(
    min_value = min(data_value, na.rm = TRUE),
    q1 = quantile(data_value, 0.25, na.rm = TRUE),
    median_value = median(data_value, na.rm = TRUE),
    q3 = quantile(data_value, 0.75, na.rm = TRUE),
    max_value = max(data_value, na.rm = TRUE)
  )|>
  knitr::kable()
```

```{r}
air_final|>
  ggplot(aes(x = data_value, fill = name)) +
  geom_histogram(binwidth = 1, position = "dodge", alpha = 0.7) +
  facet_wrap(~ name, scales = "free") +
  labs(
       x = "Data Value",
       y = "Frequency",
       fill = "Pollutant") +
  theme_minimal()
```

## Evaluation of data particles by Neighborhood

### Fine Particles by Neighborhoods

Comparison of fine particles by neighborhood 

```{r}
air_particles =
  air_final |>
  filter(name == "Fine particles (PM 2.5)")

air_particles |>
  filter(time_period == "Annual Average 2019") |>
  ggplot(aes(x = reorder(geo_place_name, -data_value), y = data_value)) +
  geom_bar(stat = "identity", fill = "skyblue")  +
  labs(
       x = "Neighborhood",
       y = "Mean Fine particles (mcg/m3)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))
```

Summary statistics of fine particles by Neighborhood

```{r}
air_final |>
  filter(name == "Fine particles (PM 2.5)") |>
  group_by(geo_place_name) |>
  summarise(
    min_value = min(data_value, na.rm = TRUE),
    q1 = quantile(data_value, 0.25, na.rm = TRUE),
    median_value = median(data_value, na.rm = TRUE),
    q3 = quantile(data_value, 0.75, na.rm = TRUE),
    max_value = max(data_value, na.rm = TRUE)
  )|>
  knitr::kable()
```

### Nitrogen Dioxide (NO2) by Neighborhoods

Comparison of NO2 by neighborhood 

```{r}
air_nitrogen =
  air_final |>
  filter(name == "Nitrogen dioxide (NO2)")

air_nitrogen |>
  filter(time_period == "Annual Average 2019") |>
  ggplot(aes(x = reorder(geo_place_name, -data_value), y = data_value)) +
  geom_bar(stat = "identity", fill = "skyblue")  +
  labs(x = "Neighborhood",
       y = "Mean Nitrogen dioxide (ppb)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))
```

Summary statistics of NO2 by Neighborhood

```{r}
air_final |>
  filter(name == "Nitrogen dioxide (NO2)") |>
  group_by(geo_place_name) |>
  summarise(
    min_value = min(data_value, na.rm = TRUE),
    q1 = quantile(data_value, 0.25, na.rm = TRUE),
    median_value = median(data_value, na.rm = TRUE),
    q3 = quantile(data_value, 0.75, na.rm = TRUE),
    max_value = max(data_value, na.rm = TRUE)
  ) |>
  knitr::kable()
```

### Ozone (NO2) by Neighborhoods

Comparison of O3 by neighborhood 

```{r}
air_final |>
  filter(name == "Ozone (O3)") |>
  ggplot(aes(x = reorder(geo_place_name, -data_value), y = data_value)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
       x = "Neighborhood",
       y = "Mean Ozone (ppb)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))
```

Summary statistics of O3 by Neighborhood

```{r}
air_final |>
  filter(name == "Ozone (O3)") |>
  group_by(geo_place_name) |>
  summarise(
    min_value = min(data_value, na.rm = TRUE),
    q1 = quantile(data_value, 0.25, na.rm = TRUE),
    median_value = median(data_value, na.rm = TRUE),
    q3 = quantile(data_value, 0.75, na.rm = TRUE),
    max_value = max(data_value, na.rm = TRUE)
  )|>
  knitr::kable()
```








