---
title: "Social Deprivation Index"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: yeti
    code_folding: hide
css: styles.css
---
<style>
  h1 {
     font-weight: bold;
  }
</style>

------------------------------------------------------------------------

[The social deprivation index (SDI)](https://www.graham-center.org/maps-data-tools/social-deprivation-index.html) is an effort to generate a scoring system of socioeconomic factors using US census data from the American Community Survey. The final SDI is a composite measure of percent living in poverty, percent with less than 12 years of education, percent single-parent households, the percentage living in rented housing units, the percentage living in the overcrowded housing unit, percent of households without a car, and percentage nonemployed adults under 65 years of age. For more details please see the webpage link.

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

library(tidyverse)
library(ggplot2)
library(leaflet)

knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE)
```
  &nbsp;

In this webpage we will explore the SDI variable by start location and determine which summary statistic to use to coalece by neighborhood.

## Load Data

```{r load_data}
citibike_df = read_csv("../citibike_clean/citibike_clean.csv")

head(citibike_df) |>
  knitr::kable()
```


 &nbsp;
  
## Summary Statistics

```{r }
citibike_df |>
  summarise(
    mean_sdi = round(mean(start_sdi_score, na.rm = TRUE), 1),
    sd_sdi = round(sd(start_sdi_score, na.rm = TRUE), 1),
    median_sdi = round(median(start_sdi_score, na.rm = TRUE), 1),
    q1_sdi = round(quantile(start_sdi_score, 0.25, na.rm = TRUE), 1),
    q3_sdi = round(quantile(start_sdi_score, 0.75, na.rm = TRUE), 1),
    iqr_sdi = round(IQR(start_sdi_score, na.rm = TRUE), 1),
    max_sdi = round(max(start_sdi_score, na.rm = TRUE), 1),
    min_sdi = round(min(start_sdi_score, na.rm = TRUE), 1)
  )|>
  pivot_longer(
    cols = c("mean_sdi", "sd_sdi", "median_sdi", "q1_sdi", "q3_sdi", "iqr_sdi", "max_sdi", "min_sdi"),
    names_to = "Statistic",
    values_to = "Value"
  )|>
  knitr::kable()
```

Since SDI score is on the zipcode level, we next decide if mean or median is the better way to iterate across neighborhoods.

```{r mean vs median by neighborhood}
citibike_df |>
  group_by(start_uhf34_neighborhood) |>
  summarise(
    mean_sdi = mean(start_sdi_score, na.rm = TRUE),
    median_sdi = median(start_sdi_score, na.rm = TRUE),
     max_sdi = round(max(start_sdi_score, na.rm = TRUE), 1),
    min_sdi = round(min(start_sdi_score, na.rm = TRUE), 1)
  )|>
  knitr::kable()
```

Mean and median SDI scores are similar by start neighborhood. However median SDI scores tend to be lower and better capture the wide variation in score across neighborhood.

  &nbsp;
  
## Median SDI by Start Zipcode

```{sdi zip}
plot = citibike_df |>
  group_by(start_zipcode) |>
  summarize(median_SDI = median(start_sdi_score)) |>
ggplot(aes(x = reorder(start_zipcode, -median_SDI), y = median_SDI,)) +
  geom_point() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  labs(x = "Start Zipcode", y = "Median SDI")  

print(plot)
 
```

It is difficult to discern the variation in start zip codes, but we can see there is a wide range.

  &nbsp;

## Median SDI by Start Neighborhood

```{r  sdi neighborhood}
plot = citibike_df |>
  group_by(start_uhf34_neighborhood) |>
  summarize(median_SDI = median(start_sdi_score)) |>
ggplot(aes(x = reorder(start_uhf34_neighborhood, -median_SDI), y = median_SDI,)) +
  geom_point() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  labs(x = "Start Neighborhood", y = "Median SDI")  

print(plot)

```
