---
title: "Weather_import"
author: "Jesus Luevano"
date: "2023-11-28"
output: github_document
---

```{r}
library(tidyverse)
```

Goal to import weather data from NCEI NOAA for 2019, https://www.ncei.noaa.gov/metadata/geoportal/rest/metadata/item/gov.noaa.ncdc:C00861/html

Download from https://www.ncei.noaa.gov/pub/data/ghcn/daily/by_year/2019.csv.gz 

Use terminal, gunzip to open

Locally import and process, then upload smaller file for ease of use on Github

               
###Data from https://www.ncei.noaa.gov/pub/data/ghcn/daily/ghcnd-stations.txt
US1NYBX0025  40.8599  -73.8457   14.9 NY BRONX 1.5 NE  
US1NYKN0003  40.6194  -73.9859    7.9 NY BROOKLYN 2.4 SW                             
US1NYKN0025  40.6846  -73.9867    5.5 NY BROOKLYN 3.1 NW
US1NYNY0074  40.7969  -73.9330    6.1 NY NEW YORK 8.8 N 
USC00300958  40.6892  -73.9550   15.2 NY BROOKLYN                                    
USC00300961  40.8372  -73.8497   11.9 NY BRONX  

US1NYQN0026  40.7544  -73.8882   18.0 NY JACKSON HEIGHTS 0.3 WSW                     
US1NYQN0027  40.7781  -73.8106   18.0 NY QUEENS 3.4 NNW                              
US1NYQN0029  40.7482  -73.7654   24.7 NY QUEENS 2.1 NE                               
US1NYQN0036  40.6719  -73.8450    6.7 NY QUEENS 4.7 SW                               
US1NYQN0037  40.7800  -73.8103   19.8 NY QUEENS 3.7 N                                
US1NYRC0001  40.5421  -74.1476   10.7 NY STATEN ISLAND 4.5 SSE                       
US1NYRC0002  40.5674  -74.1327   21.9 NY STATEN ISLAND 1.4 SE                        
US1NYRC0016  40.6157  -74.1456   11.3 NY STATEN ISLAND 2.6 N 

USC00305796  40.5939  -73.9808    6.1 NY NY AVE V BROOKLYN 
USC00305798  40.6000  -73.9667    6.1 NY NEW YORK BENSONHURST                        
USC00305799  40.8667  -73.8833   27.1 NY NEW YORK BOTANICAL GRD                      
USC00305804  40.7333  -73.9333    3.0 NY NEW YORK LAUREL HILL                        
USC00305806  40.8500  -73.9167   54.9 NY NEW YORK UNIV ST                            
USC00305816  40.7000  -74.0167    3.0 NY NEW YORK WB CITY  

The "year".csv files contain all daily and station elements found in GHCN daily for the given year.  These 
files are updated daily for the entire period of record.

The following information serves as a definition of each field in one line of data covering one station-day. 
Each field described below is separated by a comma ( , ) and follows the order below:

ID = 11 character station identification code
YEAR/MONTH/DAY = 8 character date in YYYYMMDD format (e.g. 19860529 = May 29, 1986)
ELEMENT = 4 character indicator of element type 
DATA VALUE = 5 character data value for ELEMENT 
M-FLAG = 1 character Measurement Flag 
Q-FLAG = 1 character Quality Flag 
S-FLAG = 1 character Source Flag 
OBS-TIME = 4-character time of observation in hour-minute format (i.e. 0700 =7:00 am)

See section III of the GHCN-Daily readme.txt file (ftp://ftp.ncdc.noaa.gov/pub/data/ghcn/daily/readme.txt)
for an explanation of ELEMENT codes and their units as well as the M-FLAG, Q-FLAG and S-FLAG.

The OBS-TIME field is populated with the observation times contained in NOAA/NCEI's HOMR station history database.  


```{r}

NY_Stations <- c("US1NYBX0025","US1NYKN0003", "US1NYKN0025", "US1NYNY0074", "US1NYQN0026", "US1NYQN0027", "USC00300958", "USC00300961",
                 "US1NYQN0029", "US1NYQN0036", "US1NYQN0037", "US1NYRC0001", "US1NYRC0002", "US1NYRC0016",
                 "USC00305796","USC00305798", "USC00305799", "USC00305804", "USC00305806", "USC00305816", 
                 "US1NYRC0001", "US1NYRC0002", "US1NYRC0016")
col_names_weather <- c("siteID", "date", "element","data_value", "measurement_flag", "quality_flag", "source_flag", "obs_time")

weather.df <- read_csv(file = "/Users/JML/Desktop/2019.csv",
                       col_names = col_names_weather 
                       ) %>%
  filter(siteID %in% NY_Stations)

weather.df <- weather.df %>%
  mutate(location = case_when(
    siteID == "US1NYBX0025" | siteID == "USC00300961"  ~ "Bronx", 
    siteID == "US1NYKN0003" | siteID == "US1NYKN0025" | siteID == "USC00300958" | siteID == "USC00305796" | siteID == "USC00305798" ~ "Brooklyn",
    siteID == "US1NYNY0074" | siteID == "USC00305806" | siteID == "USC00305816" ~ "Manhattan",
    siteID == "US1NYQN0026" | siteID == "US1NYQN0027" | siteID == "US1NYQN0029" | siteID == "US1NYQN0036" | siteID == "US1NYQN0037" | siteID == "USC00305799" | siteID == "USC00305804" ~ "Queens", 
    siteID == "US1NYRC0001" | siteID == "US1NYRC0002" | siteID == "US1NYRC0016" ~ "Staten"
  ))
```

