---
title: "<span style='color: blue;'>Exploration of Variables</span>"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: yeti
    code_folding: hide
css: styles.css
---

<style>
  h1 {
     font-weight: bold;
  }
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

library(tidyverse)
library(ggplot2)
library(leaflet)

knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE)
```

------------------------------------------------------------------------

# Socal Deprivation Index Exploration

 &nbsp;
 
[The social deprivation index (SDI)](https://www.graham-center.org/maps-data-tools/social-deprivation-index.html) is an effort to generate a scoring system of socioeconomic factors using US census data from the American Community Survey. The final SDI is a composite measure of percent living in poverty, percent with less than 12 years of education, percent single-parent households, the percentage living in rented housing units, the percentage living in the overcrowded housing unit, percent of households without a car, and percentage nonemployed adults under 65 years of age. For more details please see the webpage link.

In this webpage we will explore the SDI variable by start location and determine which summary statistic to use to coalesce by neighborhood.

  &nbsp;

### Load Data

```{r load SDI data}
citibike_df = read_csv("./citibike_clean/citibike_clean.csv")

head(citibike_df) |>
  knitr::kable()
```

 &nbsp;
  
### Summary Statistics

```{r }
citibike_df |>
  summarise(
    mean_sdi = round(mean(start_sdi_score, na.rm = TRUE), 1),
    sd_sdi = round(sd(start_sdi_score, na.rm = TRUE), 1),
    median_sdi = round(median(start_sdi_score, na.rm = TRUE), 1),
    q1_sdi = round(quantile(start_sdi_score, 0.25, na.rm = TRUE), 1),
    q3_sdi = round(quantile(start_sdi_score, 0.75, na.rm = TRUE), 1),
    iqr_sdi = round(IQR(start_sdi_score, na.rm = TRUE), 1),
    max_sdi = round(max(start_sdi_score, na.rm = TRUE), 1),
    min_sdi = round(min(start_sdi_score, na.rm = TRUE), 1)
  )|>
  pivot_longer(
    cols = c("mean_sdi", "sd_sdi", "median_sdi", "q1_sdi", "q3_sdi", "iqr_sdi", "max_sdi", "min_sdi"),
    names_to = "Statistic",
    values_to = "Value"
  )|>
  knitr::kable()
```

Since SDI score is on the zipcode level, we next decide if mean or median is the better way to iterate across neighborhoods.

```{r mean vs median by neighborhood}
citibike_df |>
  group_by(start_uhf34_neighborhood) |>
  summarise(
    mean_sdi = mean(start_sdi_score, na.rm = TRUE),
    median_sdi = median(start_sdi_score, na.rm = TRUE),
     max_sdi = round(max(start_sdi_score, na.rm = TRUE), 1),
    min_sdi = round(min(start_sdi_score, na.rm = TRUE), 1)
  )|>
  knitr::kable()
```

Mean and median SDI scores are similar by start neighborhood. However median SDI scores tend to be lower and better capture the wide variation in score across neighborhood.

  &nbsp;
  
### Median SDI by Start Zipcode

```{r sdi zip}
plot = citibike_df |>
  group_by(start_zipcode) |>
  summarize(median_SDI = median(start_sdi_score)) |>
ggplot(aes(x = reorder(start_zipcode, -median_SDI), y = median_SDI,)) +
  geom_point() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  labs(x = "Start Zipcode", y = "Median SDI")  

print(plot)
 
```

It is difficult to discern the variation in start zip codes, but we can see there is a wide range.

  &nbsp;

### Median SDI by Start Neighborhood

```{r  sdi neighborhood}
plot = citibike_df |>
  group_by(start_uhf34_neighborhood) |>
  summarize(median_SDI = median(start_sdi_score)) |>
ggplot(aes(x = reorder(start_uhf34_neighborhood, -median_SDI), y = median_SDI,)) +
  geom_point() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  labs(x = "Start Neighborhood", y = "Median SDI")  

print(plot)

```

------------------------------------------------------------------------

# Air Quality Dataset Exploration

 &nbsp;
 
The [air quality index (AQ)](https://data.cityofnewyork.us/Environment/Air-Quality/c3uy-2p5r) originated from NYC Open Data, contributed by the Department of Hygiene and Mental Health and include air quality indexes measured across boroughs in NYC. An exploration of AQ is below.

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

library(tidyverse)
library(ggplot2)
library(leaflet)

knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE)
```

  &nbsp;

## Load Data

```{r}
#import data
air_quality_df = read_csv("./data/air_quality/Air_Quality_20231126.csv") |>
  janitor::clean_names() |>
  mutate(
    start_date = mdy(start_date),
    year = year(start_date)
  ) |>
  filter(year == "2019")

joined_uhf_34_42 =  read_csv("./data/geocoding/joined_uhf_34_42.csv")

# add zip codes
air_geo =
  air_quality_df |>
  mutate(
    uhf34 = case_when(geo_type_name == "UHF34" ~ "1")
  )

#break up by geocoding system, add zipcodes

# uhf34
air_zip_neighborhoods_34 =
  air_geo |>
  filter(uhf34 == "1") |>
  select(-uhf34, - message)

air_zip_neighborhoods_34 =
  left_join(air_zip_neighborhoods_34, y = (joined_uhf_34_42),
            by = join_by("geo_join_id" == "uhf34")) 

air_final =
  air_zip_neighborhoods_34 |>
  select(-uhf34_neighborhood) |>
filter(!duplicated(unique_id))


head(air_final) |>
  knitr::kable()

```
 &nbsp;
  
## Summary Statistics  Across 2019

Average Air Quality Index Across Time Periods in 2019

```{r}
annual_aq = air_final |>
  filter(time_period == "Annual Average 2019") 

annual_aq|>
  ggplot(aes(x = time_period, y = data_value)) +
  geom_boxplot(fill = "skyblue", color = "black", alpha = 0.7) +
  labs(
       x = "Time Period",
       y = "Air Quality Index") +
  theme_minimal()

# Calculate statistics
midpoint <- median(annual_aq$data_value)
low_quantile <- quantile(annual_aq$data_value, 0.25)
top_quantile <- quantile(annual_aq$data_value, 0.75)

```

Summary of data particles: Fine particles (PM 2.5), Nitrogen dioxide (NO2), and Ozone (O3)

```{r}
air_final |>
  group_by(name) |>
  summarise(
    min_value = min(data_value, na.rm = TRUE),
    q1 = quantile(data_value, 0.25, na.rm = TRUE),
    median_value = median(data_value, na.rm = TRUE),
    q3 = quantile(data_value, 0.75, na.rm = TRUE),
    max_value = max(data_value, na.rm = TRUE)
  )|>
  knitr::kable()
```

```{r}
air_final|>
  ggplot(aes(x = data_value, fill = name)) +
  geom_histogram(binwidth = 1, position = "dodge", alpha = 0.7) +
  facet_wrap(~ name, scales = "free") +
  labs(
       x = "Data Value",
       y = "Frequency",
       fill = "Pollutant") +
  theme_minimal()
```

## Evaluation of data particles by Neighborhood

 &nbsp;

### Fine Particles by Neighborhoods

Comparison of fine particles by neighborhood 

```{r}
air_particles =
  air_final |>
  filter(name == "Fine particles (PM 2.5)")

air_particles |>
  filter(time_period == "Annual Average 2019") |>
  ggplot(aes(x = reorder(geo_place_name, -data_value), y = data_value)) +
  geom_bar(stat = "identity", fill = "skyblue")  +
  labs(
       x = "Neighborhood",
       y = "Mean Fine particles (mcg/m3)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))
```

Summary statistics of fine particles by Neighborhood

```{r}
air_final |>
  filter(name == "Fine particles (PM 2.5)") |>
  group_by(geo_place_name) |>
  summarise(
    min_value = min(data_value, na.rm = TRUE),
    q1 = quantile(data_value, 0.25, na.rm = TRUE),
    median_value = median(data_value, na.rm = TRUE),
    q3 = quantile(data_value, 0.75, na.rm = TRUE),
    max_value = max(data_value, na.rm = TRUE)
  )|>
  knitr::kable()
```

 &nbsp;
 
### Nitrogen Dioxide (NO2) by Neighborhoods

Comparison of NO2 by neighborhood 

```{r}
air_nitrogen =
  air_final |>
  filter(name == "Nitrogen dioxide (NO2)")

air_nitrogen |>
  filter(time_period == "Annual Average 2019") |>
  ggplot(aes(x = reorder(geo_place_name, -data_value), y = data_value)) +
  geom_bar(stat = "identity", fill = "skyblue")  +
  labs(x = "Neighborhood",
       y = "Mean Nitrogen dioxide (ppb)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))
```

Summary statistics of NO2 by Neighborhood

```{r}
air_final |>
  filter(name == "Nitrogen dioxide (NO2)") |>
  group_by(geo_place_name) |>
  summarise(
    min_value = min(data_value, na.rm = TRUE),
    q1 = quantile(data_value, 0.25, na.rm = TRUE),
    median_value = median(data_value, na.rm = TRUE),
    q3 = quantile(data_value, 0.75, na.rm = TRUE),
    max_value = max(data_value, na.rm = TRUE)
  ) |>
  knitr::kable()
```

 &nbsp;
 
### Ozone (NO2) by Neighborhoods

Comparison of O3 by neighborhood 

```{r}
air_final |>
  filter(name == "Ozone (O3)") |>
  ggplot(aes(x = reorder(geo_place_name, -data_value), y = data_value)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
       x = "Neighborhood",
       y = "Mean Ozone (ppb)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))
```

Summary statistics of O3 by Neighborhood

```{r}
air_final |>
  filter(name == "Ozone (O3)") |>
  group_by(geo_place_name) |>
  summarise(
    min_value = min(data_value, na.rm = TRUE),
    q1 = quantile(data_value, 0.25, na.rm = TRUE),
    median_value = median(data_value, na.rm = TRUE),
    q3 = quantile(data_value, 0.75, na.rm = TRUE),
    max_value = max(data_value, na.rm = TRUE)
  )|>
  knitr::kable()
```

------------------------------------------------------------------------

 &nbsp;
 
# Overweight Exploration

 &nbsp;

The [overweight data](https://a816-dohbesp.nyc.gov/IndicatorPublic/Subtopic.aspx?theme_code=2,3&subtopic_id=113) contains publicly available data from NYC regarding the percent of people who are overweight per area. An exploration of this dataset is below.

## Load Data

```{r load_data}

overweight_data = citibike_df |>
  select(start_station_latitude,
         start_station_longitude,
         start_uhf34_neighborhood,
         start_zipcode,
         end_station_latitude,
         end_station_longitude,
         end_uhf34_neighborhood,
         start_borough, end_borough,
         end_zipcode,
         start_percent_overweight, 
         end_percent_overweight) |>
  mutate(neighborhood = coalesce(end_uhf34_neighborhood, start_uhf34_neighborhood),
         lat = coalesce(end_station_latitude, start_station_latitude),
         long = coalesce(end_station_longitude, start_station_longitude),
         borough = coalesce(start_borough, end_borough),
         zipcode = coalesce(start_zipcode, end_zipcode),
         overweight = coalesce(start_percent_overweight, end_percent_overweight)) |>
  unique()


head(overweight_data) |>
  knitr::kable()
```


## Summary Statistics

```{r stats}
overweight_data |>  
  summarize(
    mean = mean(overweight, na.rm = TRUE),
    min = min(overweight, na.rm = TRUE),
    max = max(overweight, na.rm = TRUE),
    median = median(overweight, na.rm = TRUE),
    std = sd(overweight, na.rm = TRUE)
  ) |>
  knitr::kable()


```


## Percent by Zipcode
```{r zip}
percent_obese =
overweight_data |>
ggplot(aes(x = reorder(zipcode, -overweight), y = overweight,)) +
  geom_point() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  labs(title = "Percent of adults classified as overweight or obese, by area", x = "Location", y = "Percent")  

print(percent_obese)


```

## Percent by Neighborhood

```{r uhf}
percent_obese =
overweight_data |>
  group_by(neighborhood) |>
  summarize(overweight = mean(overweight)) |>
ggplot(aes(x = reorder(neighborhood, -overweight), y = overweight,)) +
  geom_point() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  labs(title = "Percent of adults classified as overweight or obese, by area", x = "Location", y = "Percent")  

print(percent_obese)


```


```{r}
# Create a leaflet map

# Select the necessary columns
map_data <- overweight_data %>%
  select(lat, long, overweight)

map <- leaflet(data = map_data) %>%
  addTiles()  # Add map tiles (you can use different tile providers)

# Add color-coded circles based on overweight percentage
map <- map %>%
  addCircleMarkers(
    radius = 5,  # Adjust the circle size as needed
    fillColor = ~colorFactor("Blues", map_data$overweight)(overweight),
    color = "black",
    fillOpacity = 0.7,
    popup = ~paste("Overweight Percentage:", overweight, "%"),
    label = ~paste("Overweight Percentage:", overweight, "%")
  )

# Display the map
map

```









