---
title: "From Pedals to Patterns: A P8105 Citibike Data Analysis Over 12 months of 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: yeti
---

# About our Project:
```{r library, include=FALSE}
library(tidyverse)
library(shiny)
library(shinyjs)
shinyjs::useShinyjs()
```


## Screencast 

<link here to embed video>

## Team Members
  - Laura O'Carroll, lgo2107<br>
  - Courtney Diamond, cjd2195<br>
  - Hyun Jin Jung, hj2660<br>
  - Jesus Luevano, jl5934<br>
  - Kayla Schiffer-Kane, khs2138<br>
  - Haley Zylberberg, hmz2105<br>
  
  
## Motivation
Increasing transportation costs, congested streets, and subway delays across New York City are pressuring residents to find cheaper and faster means of navigating the Big Apple. CitiBikes have had an impact in increasing access to biking as a viable alternative to other means of transportation. There is an added health benefit to the physical activity of biking. Bike usage also confers health related benefits on air quality. The utilization of CitiBikes, however, is dependent on factors such as availability of bike stations, bike lanes, and the cost of bike rides and membership. 

We, as young professionals/students in the city, want to know if turning to CitiBike is a viable alternative. We also want to explore if there is disproportionate availability and utilization of CitiBikes across neighborhoods with varied social deprivation index that can exacerbate disparities in health status. To that end, we will explore CitiBike usage, aiming to understand usage patterns related to location within NYC, and its relation to social deprivation index, air quality, and health outcomes.


## Sources of Data

[**CitiBike Data**](https://s3.amazonaws.com/tripdata/index.html) We will use the publicly available CitiBike. These datasets contain monthly logs of trips and record data elements such as the ride ID, the user's status (subscriber or casual user), the trip start location and longitude/latitude coordinates, the trip end location and latitude/longitude coordinates, and the trip start/stop times and duration.

* The available data also vary by year. For example, 2023 contains data on mechanical versus electric break usage and whether the rider was a CitiBike member or not. Pending initial exploration of completion, we will determine feasibility of incorporating these different variables in our analysis. 

[**Air Quality Data**](https://data.cityofnewyork.us/Environment/Air-Quality/c3uy-2p5r) Additionally, we plan to investigate how weather trends impact Citibike ridership in the area using data from NYC Open Data, contributed by the Department of Hygiene and Mental Health. These data include the air quality indexes measured across boroughs in NYC.

[**Social Deprivation Index**](https://www.graham-center.org/maps-data-tools/social-deprivation-index.html) Social deprivation index (SDI) is an effort in generating composite measures of socioeconomic factors across NYC by census tract. We will use these datasets to explore CitiBike availability and usage patterns across neighborhoods.

[**Overweight Data**](https://a816-dohbesp.nyc.gov/IndicatorPublic/Subtopic.aspx?theme_code=2,3&subtopic_id=113) To provide an insight into health outcomes associated with factors such as physical acitvity aand socioeconomic factors, we will use NYC's publicly available overweight health outcomes data, which is.  

[**Weather Data**](https://www.ncei.noaa.gov/metadata/geoportal/rest/metadata/item/gov.noaa.ncdc:C00861/html) Lastly, for temperature-related trends, we will use data from the Global Historical Climatology Network - Daily (GHCN-Daily) dataset published by the National Centers for Environmental Information (NCEI). These data reflect those used in the rnoaa package, and include the weather station identifier, the date, latitude/longitude, maximum temperature, minimum temperature, and precipitation. 


# Dataset exploration

## Exploring Citibikes 

We utilized the datasets from the above Citibank link, specifically downloading monthly datasets for NYC data from December 2018 - January 2020 (to encompass the entirety or rides within NYC in 2019, and to include those that began Dec 31st, 2018 into Jan 1st 2019, and that began Dec 31st, 2019 into Jan 1st 2020). 

Data was imported and consolidated into one dataset, with filtering done to remove rides that were <5 minutes or >24hours in length. 

```{r, Citibike dataset}

```

In addition we imported each of the sub-datasets:

```{r, air quality dataset}

```


```{r, SDI dataset}

```

```{r, weight dataset}

```


```{r, weather_import, eval=FALSE}
#This file was >1GB, so downlaoded and cleaned locally, uploaded to GitHub after filtering

NY_Stations <- c("US1NYBX0025","US1NYKN0003", "US1NYKN0025", "US1NYNY0074", "US1NYQN0026", "US1NYQN0027", "USC00300958", "USC00300961",
                 "US1NYQN0029", "US1NYQN0036", "US1NYQN0037", "US1NYRC0001", "US1NYRC0002", "US1NYRC0016",
                 "USC00305796","USC00305798", "USC00305799", "USC00305804", "USC00305806", "USC00305816", 
                 "US1NYRC0001", "US1NYRC0002", "US1NYRC0016")
col_names_weather <- c("siteID", "date", "element","data_value", "measurement_flag", "quality_flag", "source_flag", "obs_time")

weather.df <- read_csv(file = "/Users/JML/Desktop/2019.csv",
                       col_names = col_names_weather 
                       ) %>%
  filter(siteID %in% NY_Stations)

weather.df <- weather.df %>%
  mutate(location = case_when(
    siteID == "US1NYBX0025" | siteID == "USC00300961"  ~ "Bronx", 
    siteID == "US1NYKN0003" | siteID == "US1NYKN0025" | siteID == "USC00300958" | siteID == "USC00305796" | siteID == "USC00305798" ~ "Brooklyn",
    siteID == "US1NYNY0074" | siteID == "USC00305806" | siteID == "USC00305816" ~ "Manhattan",
    siteID == "US1NYQN0026" | siteID == "US1NYQN0027" | siteID == "US1NYQN0029" | siteID == "US1NYQN0036" | siteID == "US1NYQN0037" | siteID == "USC00305799" | siteID == "USC00305804" ~ "Queens", 
    siteID == "US1NYRC0001" | siteID == "US1NYRC0002" | siteID == "US1NYRC0016" ~ "Staten"
  ))
```


<Link to page with exploratory analyses, plots for Citibank>


## Exploring Subsets

[SDI](SDI_Exploratory.html)

[Air_Quality](SDI_Exploratory.html)

[Weight](Weight_exploratory.html)

[Weather](Weather_exploratory.html)


# Modeling outcomes

** 
