---
title: "NOAA Weather dataset exploratoration"
output: html_document
---

## Weather dataset preparation 
We sought to gather meteorological data from the National Centers for Environmental Information (NCEI), which is part of the National Oceanic and Atmospheric Administration (NOAA). 

The data was found on the [NOAA](https://www.ncei.noaa.gov/metadata/geoportal/rest/metadata/item/gov.noaa.ncdc:C00861/html) website, and we selected to downlaod the entirety of data for the year of 2019 as our starting database. The specific [file](https://www.ncei.noaa.gov/pub/data/ghcn/daily/by_year/2019.csv.gz)

We then used the terminal command `gunzip` to open the .gz file. Given the size of the file was >1GB we processed it locally with the goal to filter based on sites that were within NYC. 

We found sites by reviewing the list of all of them [sites](https://www.ncei.noaa.gov/pub/data/ghcn/daily/ghcnd-stations.txt). The NYC sites were saved in a vector `NY_Stations` to then filter the database output. 

The initial file had no column headers, so we had to create headers and saved in `col_names_weather`.


```{r weather dataframe prep, eval=FALSE}
NY_Stations <- c("US1NYBX0025","US1NYKN0003", "US1NYKN0025", "US1NYNY0074", "US1NYQN0026", "US1NYQN0027", "USC00300958", "USC00300961",
                 "US1NYQN0029", "US1NYQN0036", "US1NYQN0037", "US1NYRC0001", "US1NYRC0002", "US1NYRC0016",
                 "USC00305796","USC00305798", "USC00305799", "USC00305804", "USC00305806", "USC00305816", 
                 "US1NYRC0001", "US1NYRC0002", "US1NYRC0016")
col_names_weather <- c("siteID", "date", "element","data_value", "measurement_flag", "quality_flag", "source_flag", "obs_time")

weather.df <- read_csv(file = "/Users/JML/Desktop/2019.csv",
                       col_names = col_names_weather 
                       ) %>%
  filter(siteID %in% NY_Stations)

weather.df <- weather.df %>%
  mutate(location = case_when(
    siteID == "US1NYBX0025" | siteID == "USC00300961"  ~ "Bronx", 
    siteID == "US1NYKN0003" | siteID == "US1NYKN0025" | siteID == "USC00300958" | siteID == "USC00305796" | siteID == "USC00305798" ~ "Brooklyn",
    siteID == "US1NYNY0074" | siteID == "USC00305806" | siteID == "USC00305816" ~ "Manhattan",
    siteID == "US1NYQN0026" | siteID == "US1NYQN0027" | siteID == "US1NYQN0029" | siteID == "US1NYQN0036" | siteID == "US1NYQN0037" | siteID == "USC00305799" | siteID == "USC00305804" ~ "Queens", 
    siteID == "US1NYRC0001" | siteID == "US1NYRC0002" | siteID == "US1NYRC0016" ~ "Staten"
  ))
```

## Exploratory analyses

```{r}

```

