---
title: "Citibank Exploratoration"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

library(ggridges)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Only run the below chunk if running local, not knitting: 
```{r local_filepath, eval = FALSE}
mypath = 'C:/Users/16175/Desktop/pedalstopatterns.github.io.git'  # Replace with your actual path
path = paste0(mypath, '/exploratory')
setwd(path)
```


# Citibanks Dataset
Jan/2019 ~ Dec/2019
```{r import_citibike, echo = FALSE}
citibike <- 
 tibble(
 files = list.files("../citibike/"),
  path = str_c("../citibike/", files)
  ) |>
mutate(data = map(path, ~read_csv(.x, col_types = cols(
  'end station id' = col_double(),
   'start station id' = col_double()
  )))) |>
unnest(cols = c(data))


#tidy citibike dataset
citibike_df <- citibike |>
  janitor::clean_names() |>
  select(-files, -path) |>
  rename(trip_duration_sec = tripduration,
         start_time = starttime,
         stop_time = stoptime,
         user_type = usertype)|>
  mutate(gender = recode(gender,
                         "0" = "Unknown",
                         "1" = "Male",
                         "2" = "Female"),
        trip_duration_min = trip_duration_sec / 60,
        age = 2019 - birth_year
        ) |>
  filter(
    trip_duration_sec >= 300,
    trip_duration_sec <= 86400,
    as.Date(stop_time) >= as.Date("2019-01-01"),
    as.Date(start_time) <= as.Date("2019-12-31")
  )  |>
  select(trip_duration_sec, trip_duration_min, everything())
```


Citibike to Neighborhood

```{r citibike_neighborhood, echo = FALSE}
latlong_zip = read_csv('../data/geocoding/citibike_latlong_zip.csv')
joined_uhf_34_42 = read_csv('../data/geocoding/joined_uhf_34_42.csv')

citibike_zip =
  citibike_df |>
  left_join(latlong_zip, 
            by = c("start_station_latitude" = "latitude",
                   "start_station_longitude" = "longitude")) |>
  rename("start_zipcode" = "postcode") |>
  left_join(latlong_zip, 
            by = c("end_station_latitude" = "latitude",
                   "end_station_longitude" = "longitude")) |>
  rename("end_zipcode" = "postcode") 

citibike_zip_neighborhoods =
  citibike_zip |>
  left_join(y = (joined_uhf_34_42 |> select(zip, uhf42_neighborhood, uhf34_neighborhood)),
            by = join_by("start_zipcode" == "zip")) |> 
  rename("start_uhf34_neighborhood" = "uhf34_neighborhood",
         "start_uhf42_neighborhood" = "uhf42_neighborhood") |> 
  left_join(y = (joined_uhf_34_42 |> select(zip, uhf42_neighborhood, uhf34_neighborhood)),
            by = join_by("end_zipcode" == "zip")) |> 
  rename("end_uhf34_neighborhood" = "uhf34_neighborhood",
         "end_uhf42_neighborhood" = "uhf42_neighborhood")

head(citibike_zip_neighborhoods)
```

## Neighborhoods Analyses

* Rides per start neighborhood per month
```{r neighborhood_month}
neighborhood_month = citibike_zip_neighborhoods |>
  mutate(month = as.factor(month(start_time))) |>
  group_by(start_uhf34_neighborhood, month) |>
  summarize(n_rides = n())

neighborhood_month |> 
  ggplot(aes(x = month, y = n_rides, color = start_uhf34_neighborhood, group = start_uhf34_neighborhood)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = scales::comma_format()) +
  labs(x = "Month", y = "Number of Rides")
```


* Rides starting in each borough 

### Note (Kayla): Borough is in zipcodetable.pdf, but don't think we've included it? 
```{r neighborhood_borough}


```


* Rides ending in each borough

* Most popular landmark stations, and rides by month
```{r}
# Count the number of rides for each start station
popular_stations <- citibike_zip_neighborhoods %>%
  group_by(start_station_name) %>%
  summarise(n_rides = n()) %>%
  top_n(10, wt = n_rides)  # Select the top 10 most popular stations

# Plot the most popular landmark stations
ggplot(popular_stations, aes(x = reorder(start_station_name, n_rides), y = n_rides)) +
  geom_bar(stat = "identity") +
  labs(x = "Start Station", y = "Number of Rides", title = "Top 10 Most Popular Landmark Stations") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Count the number of rides for each month
rides_by_month <- citibike_zip_neighborhoods %>%
  mutate(month = format(start_time, "%Y-%m")) %>%
  group_by(month) %>%
  summarise(n_rides = n())

# Plot the number of rides by month
ggplot(rides_by_month, aes(x = month, y = n_rides)) +
  geom_line() +
  geom_point() +
  labs(x = "Month", y = "Number of Rides", title = "Number of Rides by Month") +
  scale_y_continuous(labels = scales::comma_format()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

* User demographics (subscriber vs customer, gender, birth years)
```{r}
# Count the number of rides for each user type
user_type_counts <- citibike_zip_neighborhoods %>%
  group_by(user_type) %>%
  summarise(n_rides = n())

# Plot the user type distribution
ggplot(user_type_counts, aes(x = user_type, y = n_rides, fill = user_type)) +
  geom_bar(stat = "identity") +
  labs(x = "User Type", y = "Number of Rides", title = "User Type Distribution")

# Count the number of rides for each gender
gender_counts <- citibike_zip_neighborhoods %>%
  group_by(gender) %>%
  summarise(n_rides = n())

# Plot the gender distribution
ggplot(gender_counts, aes(x = gender, y = n_rides, fill = gender)) +
  geom_bar(stat = "identity") +
  labs(x = "Gender", y = "Number of Rides", title = "Gender Distribution")

# Plot a histogram of birth years
ggplot(citibike_zip_neighborhoods, aes(x = birth_year)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(x = "Birth Year", y = "Number of Rides", title = "Birth Years Distribution")
```

